---
title: "ds202_hw5"
author: "Syed Hasan github:https://github.com/shasan2519/ds202_lab5"
date: "4/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=F}
accident <- read.csv("accident.csv")
person <- read.csv("person.csv")
```

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

# Part 1
```{r}
countDay <- count(accident,DAY_WEEK)
countDay
Days<- max(countDay)
Days
```
  
Explaination: Accidents are happening almost on a daily basis. 
  

```{r}

countHour <- count(accident,HOUR) 
countHour
Hours<- max(countHour)
Hours

```
  
Explaination: Accident happening at all hours.
  

```{r}
Drunk=subset(accident, accident$DRUNK_DR > 0 )
nrow(Drunk)
```
  
Answer: 8769 drunk drivers

  

# Part 2
```{r}
p2 <- person %>% filter(PER_TYP==1) %>% right_join(accident)


daily <- p2 %>% group_by(DAY_WEEK) %>% summarize(count = n())
daily$DAY_WEEK <- as.factor(daily$DAY_WEEK)
levels(daily$DAY_WEEK) <-   c("Su","M","T","W","Th","F","Sa")

h <- p2 %>% group_by(HOUR) %>% summarize(h = n())

p2$SEX <- as.factor(p2$SEX)
levels(p2$SEX) <- c("M","F","U","U")
p2$SEX[is.na(p2$SEX)] <- "U"



sex <- p2 %>% group_by(SEX) %>% summarize(gender = n())
```

```{r}
daily
daily %>% ggplot(aes(x=DAY_WEEK,y=count))+ geom_col() + labs(x="week Day",y=" accident tally")
```
  
Weekends have a higher rate of accidents
  
```{r}
h
h %>% filter(HOUR<90) %>% ggplot(aes(x=HOUR,y=h))+ geom_step() + geom_point() + labs(x="time",y="accident tally")
```
  
highers accidents during the late hours wiht lowest around Dawn time
  
```{r}
sex
sex %>%filter(SEX %in% c("M","F")) %>% ggplot(aes(x=SEX,y=gender))+ geom_col() + labs(x="sex",y="accident tally")
```
  
Femlaes are less likely to crash than males
  
    
# Part 3  
  
```{r}
library(maps)
library(mapproj)
us1 <- readxl::read_xlsx(("US_GLC.xlsx"), col_types = c("text","text","text","text","text","text","text","text","text","skip"))
ust <- readxl::read_xlsx("UST_GLC.xlsx")

us1$STATE <- as.numeric(us1$`State Code`)
us1$COUNTY <- as.numeric(us1$`County Code`)


p31 <- accident %>% left_join(us1)
head(p31)
county <- map_data('county')

p32 <- p31 %>% select(`State Name`,`County Name`,MONTH)

p3summer <- p32 %>% group_by(`State Name`,`County Name`) %>% summarize(count = n())

p3summer$region <- tolower(p3summer$`State Name`)
p3summer$subregion <- tolower(p3summer$`County Name`)

p3 <- p3summer %>%select(region, subregion, count) %>% inner_join(county)
par(mfrow=c(2,1))
p3 %>% ggplot(aes(x=long,y=lat,group=group)) + geom_path() + geom_polygon(aes(fill=count))  + scale_fill_gradient(low='green',high='black')+coord_map() + labs(x="long",y="lat",title="Choropleth Plot")

```

Major green (high accident rate) areas are all Chicago, Parts of texas, Parts of southern California
   
  
```{r}
state <- map_data('state')

p3_summer <- p32 %>% filter(MONTH %in% c(4:9)) %>% mutate(region = tolower(`State Name`)) %>% group_by(region) %>% summarize(count = n())

p3winter <- p32 %>% filter(MONTH %in% c(1:3,10:12)) %>%mutate(region = tolower(`State Name`)) %>%group_by(region) %>%summarize(count = n())

p3season <- data.frame(region=p3winter$region,diff = (p3_summer$count-p3winter$count))
p3f <- p3season %>%inner_join(state)

p3f %>% ggplot(aes(x=long,y=lat,group=group)) + geom_path() + geom_polygon(aes(fill=diff))  + scale_fill_gradient(low='green',high='black')+coord_map() + labs(x="long",y="lat")
```
  
All the state where snow is most presistent like IL, New york and midwest have high a differnce

